Sampbias, a method to evaluate geographic sampling bias in species distribution data

### 2019-12-18 manuscript.rmd
empirical results and appendix S1

---
title: "Sampbias, a method to evaluate geographic sampling bias in species distribution data"
#author: "Alexander Zizka^1,2,3^, Alexandre Antonelli^3,4,5^, Daniele Silvestro^3,4^"
bibliography: sampbias.bib
csl: ecography.csl
output: 
  bookdown::pdf_document2:
  #word_document:
    toc: false
    toc_depth: 1
    fig_caption: true
    number_sections: false
header-includes: |
    \usepackage{caption}
    \usepackage{fancyhdr}
    \pagestyle{fancy}
    \fancyhead[L]{sampbias accessibility bias}
    \fancyhead[R]{\thepage}
    \fancyfoot{}
    \usepackage{float}
    \floatplacement{figure}{H}
    \usepackage{lineno}
    \linenumbers
    \usepackage[url=false]{biblatex}
linestretch: 2
fontsize: 12pt
always_allow_html: yes
---

@@

[**Daniele some text here on the projection of bias through time, and what the plots exactly show**]

@@

To exemplify the use and output of sampbias, we downloaded the occurrence records of all mammals available from the Indonesian island of Borneo [@gbifdoi2016], and ran *sampbias* as shown in the example code below, to test the biasing effect of airports, cities and roads in the dataset. The example dataset is provided with the *sampbias*. On a 0.05 degree scale (~ 5 km) we found 
(Fig. \ref{fig:empirical}).

@@

#project in space
proj <- project_bias(example.out)
map_bias(proj)

```

# Data accessibility
The software presented here is available under a GPL-3 license. The *sampbias* R package is available https://github.com/azizka/sampbias. The R package includes an example dataset as well as vignetts detailing the use of the R package (Appendix S1),  and possibly warnings produced by the package (Appendix S2).

@@

Appendix S1 - Tutorial running sampbias in R

Appendix S2 - Possible warnings and their solutions

# References

>>>>

---
title: "Sampbias, a method to evaluate geographic sampling bias in species distribution data"
#author: "Alexander Zizka^1,2,3^, Alexandre Antonelli^3,4,5^, Daniele Silvestro^3,4^"
bibliography: sampbias.bib
csl: ecography.csl
output: 
  bookdown::pdf_document2:
  #word_document:
    toc: false
    toc_depth: 1
    fig_caption: true
    number_sections: false
header-includes: |
    \usepackage{caption}
    \usepackage{fancyhdr}
    \pagestyle{fancy}
    \fancyhead[L]{sampbias accessibility bias}
    \fancyhead[R]{\thepage}
    \fancyfoot{}
    \usepackage{float}
    \floatplacement{figure}{H}
    \usepackage{lineno}
    \linenumbers
    \usepackage[url=false]{biblatex}
    \usepackage{setspace}
linestretch: 2
fontsize: 12pt
always_allow_html: yes
---

@@

[**Daniele some text here on the projection of bias through space, and what the plots exactly show**]

@@

To exemplify the use and output of *sampbias*, we downloaded the occurrence records of all mammals available from the Indonesian island of Borneo [n = 6,262, @gbifdoi2016], and ran *sampbias* using the default gazetteers as shown in the example code below, to test the biasing effect of the main airports, cities and roads in the dataset. The example dataset is provided with the *sampbias*. We found a strong effect of cities on sampling intensity, a moderate effect of roads and airports and no effect of rivers (Fig. \ref{fig:empirical}A). All models predict a low number of collection records in the centre of Borneo, which reflects the original data, and where accessibility means are low (Figure S1 in Appendix S1). We ran this example to illustrate the use of *sampbias* For detailed analyses or a smaller geographic scale, higher resolution gazetteers, including smaller roads and rivers  and a higher spatial resolution would be desirable. Results might change with increasing resolution, since roads and rivers might have a stronger effect on higher resolutions (facilitating most the access to their immediate vicinity), whereas cities and airports might have an effect on the larger scale (facilitating access to a larger area).

@@

#project in space
proj <- project_bias(example.out)
map_bias(proj)
```

# Data accessibility
The software presented here is available under a GPL-3 license. The *sampbias* R package is available https://github.com/azizka/sampbias. The R package includes an example dataset as well as vignettes detailing the use of the R package (Appendix S1),  and possibly warnings produced by the package (Appendix S2).

@@

Appendix S1 - Supplementary Figure

Appendix S2 - Tutorial running sampbias in R

Appendix S3 - Possible warnings and their solutions

\newpage{}
# References
